---
title: "Module 3 — Tables, Figures & Layout"
author: "Your Name"
date: today
format:
  typst:
    papersize: a4
    margin:
      x: 2.5cm
      y: 2.5cm
    fontsize: 11pt
    number-sections: true
fig-format: png
fig-dpi: 300
---

```{r}
#| label: setup
#| include: false
library(knitr)
library(ggplot2)
```

# Tables with `kable()`

`knitr::kable()` output is converted into a native Typst table automatically.
Captions and cross-reference labels work exactly as in other Quarto formats:

```{r}
#| label: tbl-mtcars
#| tbl-cap: "Fuel efficiency of selected cars (mtcars dataset)."
top5 <- mtcars[order(-mtcars$mpg), c("mpg", "cyl", "wt", "hp")][1:5, ]
kable(top5, digits = 1)
```

See @tbl-mtcars for the five most efficient cars — that reference number was
generated automatically because the chunk label starts with `tbl-`.

# Figures

Plots are embedded as images. Use `fig-cap` for captions and `fig-` labels
for cross-referencing:

```{r}
#| label: fig-mpg-wt
#| fig-cap: "Heavier cars use more fuel."
#| fig-width: 6
#| fig-height: 3.5
ggplot(mtcars, aes(wt, mpg)) +
  geom_point(colour = "#1f6f8b", size = 3) +
  geom_smooth(method = "lm", se = FALSE, colour = "grey40",
              formula = y ~ x) +
  labs(x = "Weight (1000 lbs)", y = "Miles per gallon") +
  theme_minimal(base_size = 12)
```

@fig-mpg-wt shows the classic weight–efficiency relationship.

# Side-by-side layout

Quarto's layout attributes work in Typst. Two plots in one row:

::: {layout-ncol=2}
```{r}
#| label: fig-iris-a
#| fig-cap: "Sepal dimensions."
#| fig-height: 4
ggplot(iris, aes(Sepal.Length, Sepal.Width, colour = Species)) +
  geom_point() + theme_minimal() + theme(legend.position = "bottom")
```

```{r}
#| label: fig-iris-b
#| fig-cap: "Petal dimensions."
#| fig-height: 4
ggplot(iris, aes(Petal.Length, Petal.Width, colour = Species)) +
  geom_point() + theme_minimal() + theme(legend.position = "bottom")
```
:::

# Styling tables with Typst show rules

You can restyle *all* tables in one stroke with a raw Typst show rule.
This one adds a subtle striped background:

```{=typst}
#set table(
  stroke: 0.5pt + rgb("#b0c4cc"),
  fill: (_, y) => if y > 0 and calc.odd(y) { rgb("#f2f7f9") },
)
#show table.cell.where(y: 0): strong
```

```{r}
#| label: tbl-tooth
#| tbl-cap: "Tooth growth by supplement (first rows)."
kable(head(ToothGrowth, 6))
```

(If a show rule ever misbehaves, remember `keep-typ: true` lets you inspect
exactly what Typst received.)

# Full-width elements in narrow layouts

If your document uses `columns: 2`, a raw Typst `#place` or Quarto's
`.column-page` div can break out. Simplest tool: Typst's page-spanning
figure placement:

````markdown
```{=typst}
#figure(placement: top, ...)
```
````

We keep this document single-column, but try it in Exercise 3.4.

# Exercises

## 🏋️ Exercise 3.1 — A referenced table

Create a table of the first 6 rows of `airquality` (columns `Ozone`,
`Temp`, `Wind` only), give it the caption "New York air quality, May 1973",
label it `tbl-air`, and reference it from a sentence.

## 🏋️ Exercise 3.2 — A referenced figure

Plot `Temp` against `Ozone` from `airquality` with a smoother, caption it,
label it `fig-air`, and reference it in text.

## 🏋️ Exercise 3.3 — Three-across layout

Use `::: {layout-ncol=3}` to place three histograms of `iris` measurements
side by side.

## 🏋️ Exercise 3.4 — Two-column report page

Copy this file, set `columns: 2` in the YAML, and re-render. Observe what
happens to wide tables — then shrink the mtcars table (fewer columns, or
`fontsize` via a Typst set rule) until it fits a column.

::: {.callout-tip}
Solutions in `solutions/03-solutions.qmd`.
:::
