---
title: "Module 3 — Solutions"
format:
  typst:
    papersize: a4
    margin:
      x: 2.5cm
      y: 2.5cm
    number-sections: true
fig-format: png
fig-dpi: 300
---

```{r}
#| label: setup
#| include: false
library(knitr)
library(ggplot2)
```

# Exercise 3.1 — A referenced table

```{r}
#| label: tbl-air
#| tbl-cap: "New York air quality, May 1973."
kable(head(airquality[airquality$Month == 5, c("Ozone", "Temp", "Wind")]))
```

As @tbl-air shows, early-May ozone readings were modest.

# Exercise 3.2 — A referenced figure

```{r}
#| label: fig-air
#| fig-cap: "Ozone rises sharply with temperature."
#| fig-width: 6
#| fig-height: 3.5
#| warning: false
ggplot(airquality, aes(Temp, Ozone)) +
  geom_point(colour = "#1f6f8b", alpha = 0.7) +
  geom_smooth(method = "loess", se = FALSE, colour = "grey40",
              formula = y ~ x) +
  labs(x = "Temperature (°F)", y = "Ozone (ppb)") +
  theme_minimal(base_size = 12)
```

@fig-air suggests a strongly non-linear relationship.

# Exercise 3.3 — Three-across layout

::: {layout-ncol=3}
```{r}
#| fig-height: 4
hist(iris$Sepal.Length, main = "Sepal length", col = "#1f6f8b",
     border = "white", xlab = "cm")
```

```{r}
#| fig-height: 4
hist(iris$Sepal.Width, main = "Sepal width", col = "#1f6f8b",
     border = "white", xlab = "cm")
```

```{r}
#| fig-height: 4
hist(iris$Petal.Length, main = "Petal length", col = "#1f6f8b",
     border = "white", xlab = "cm")
```
:::

# Exercise 3.4 — Two-column notes

Setting `columns: 2` in the YAML narrows every text block to half width.
Wide tables overflow. Two fixes that work well together:

1. Show fewer columns: `mtcars[, c("mpg", "cyl", "wt")]`.
2. Shrink table text with a raw Typst rule before the table:

```{=typst}
#show table: set text(size: 8pt)
```

```{r}
#| label: tbl-narrow
#| tbl-cap: "A table slimmed down to fit a narrow column."
kable(head(mtcars[, c("mpg", "cyl", "wt")]), digits = 1)
```
