---
title: "New York Air Quality, Summer 1973"
subtitle: "Capstone worked solution"
author: "Jane Trainer"
date: today
format:
  typst:
    papersize: a4
    margin:
      x: 2.5cm
      y: 2.5cm
    fontsize: 10pt
    mainfont: "Libertinus Serif"
    toc: true
    number-sections: true
fig-format: png
fig-dpi: 300
---

```{=typst}
#show heading.where(level: 1): set text(fill: rgb("#1f6f8b"))
#set par(justify: true)
```

```{r}
#| label: setup
#| include: false
library(knitr)
library(ggplot2)
aq <- na.omit(airquality)
aq$MonthName <- factor(month.name[aq$Month], levels = month.name[5:9])
```

# Key findings

```{=typst}
#block(
  fill: rgb("#eef6f8"),
  stroke: (left: 3pt + rgb("#1f6f8b")),
  inset: 10pt,
  radius: 4pt,
  width: 100%,
)[
  *In brief.* Ozone pollution peaked in July and August, tracks temperature
  strongly and non-linearly, and is moderated by wind. August 25 recorded
  the summer's worst reading.
]
```

```{r}
#| label: headline-stat
#| output: asis
#| echo: false
i <- which.max(aq$Ozone)
cat("```{=typst}\n",
    "#align(center)[",
    "#block(fill: rgb(\"#fff4e6\"), inset: 12pt, radius: 6pt)[",
    "#text(size: 20pt, weight: \"bold\")[", aq$Ozone[i], " ppb]\n\n",
    "Peak ozone, recorded on ", aq$MonthName[i], " ", aq$Day[i],
    "]]\n```", sep = "")
```

# Monthly summary

```{r}
#| label: tbl-monthly
#| tbl-cap: "Monthly means of ozone, temperature and wind speed."
monthly <- aggregate(cbind(Ozone, Temp, Wind) ~ MonthName, aq, mean)
names(monthly)[1] <- "Month"
kable(monthly, digits = 1)
```

@tbl-monthly shows the mid-summer peak: mean ozone in July and August was
roughly triple the May level, coinciding with the warmest temperatures and
the calmest winds of the season.

# Ozone and temperature

```{r}
#| label: fig-ozone-temp
#| fig-cap: "Ozone concentration versus daily maximum temperature."
#| fig-width: 6
#| fig-height: 3.5
ggplot(aq, aes(Temp, Ozone)) +
  geom_point(aes(colour = MonthName), alpha = 0.8) +
  geom_smooth(method = "loess", se = FALSE, colour = "grey30",
              formula = y ~ x) +
  labs(x = "Temperature (°F)", y = "Ozone (ppb)", colour = NULL) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")
```

@fig-ozone-temp shows ozone is flat below about 80 °F, then climbs steeply —
consistent with photochemical ozone formation accelerating in heat.

# Two moderating factors

::: {layout-ncol=2}
```{r}
#| label: fig-wind
#| fig-cap: "Wind disperses ozone."
#| fig-height: 4
ggplot(aq, aes(Wind, Ozone)) +
  geom_point(colour = "#1f6f8b", alpha = 0.7) +
  geom_smooth(method = "loess", se = FALSE, colour = "grey30",
              formula = y ~ x) +
  labs(x = "Wind (mph)", y = "Ozone (ppb)") +
  theme_minimal(base_size = 11)
```

```{r}
#| label: fig-solar
#| fig-cap: "Sunlight fuels ozone formation."
#| fig-height: 4
ggplot(aq, aes(Solar.R, Ozone)) +
  geom_point(colour = "#c0392b", alpha = 0.7) +
  geom_smooth(method = "loess", se = FALSE, colour = "grey30",
              formula = y ~ x) +
  labs(x = "Solar radiation (lang)", y = "Ozone (ppb)") +
  theme_minimal(base_size = 11)
```
:::

# Methods note

Analysis uses the base-R `airquality` dataset (153 daily observations,
May–September 1973, Roosevelt Island / LaGuardia). Rows with missing ozone
or solar readings were dropped (`na.omit`), leaving `r nrow(aq)` days.
